### 2020-04-02 material/analysis/ioi_onset.Rmd
this is only for onsets
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'IOIs Analysis for Non-Expressive performances (ONSET)'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for a tempo (IOIs).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
df_ioi <- fread("./../preprocessor/filtered/data_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

# assign RowNr
df_ioi$RowNr <- rep(1:72, nrow(df_ioi)/72)
```

```{r normIOIs, include = FALSE}
# calculate normIOIs
df_ioi$IOI <- diff(c(0, df_ioi$TimeStamp))

# convert bpm to ms
df_ioi[Tempo == 120]$Tempo <- 250
df_ioi[Tempo == 110]$Tempo <- 273
df_ioi[Tempo == 100]$Tempo <- 300

# normalise IOIs
df_ioi$normIOI <- df_ioi$IOI/df_ioi$Tempo

# remove the first note
df_ioi <- df_ioi[RowNr != 1]

# assign Interval
df_ioi$Interval <- rep(1:71, nrow(df_ioi)/71)

# remove irrelevant intervals
ls_irrelevant <- list(c(25:28), c(32:35), c(51:52), c(56:57), c(65:71))
df_ioi$EighthNote <- "Yes"

for (phrase in 1:length(ls_irrelevant)){
  for (note in 1:length(ls_irrelevant[[phrase]])){
    df_ioi[Interval == ls_irrelevant[[phrase]][note]]$EighthNote <- "No"
  }
}
```

# IOI
## 1. All - individual
- Average normIOIs for each participant.

```{r ioi, echo = FALSE}
# For each individual
ioi <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), by = .(SubNr)]
print(ioi)
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the ideal tempo

```{r ioi_bar,  echo = FALSE}
p_ioi <- ggplot(data = ioi, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  labs(y = "Normalised IOI", subtitle = "IOI: Non-Expression")
p_ioi
```

## 2. All - group
- Average normIOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi_all, echo = FALSE}
# Group mean
ioi_all <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI))]
print(ioi_all)
```

## Sequence plot
```{r seq, fig.width = 10, echo = FALSE}
# For each individual
ioi_seq <- df_ioi[EighthNote == "Yes", .(N = length(normIOI), Mean = mean(normIOI), SD = sd(normIOI)), .(SubNr, Interval)]

p_ioi_seq <- ggplot(data = ioi_seq, aes(x = Interval, y = Mean, colour = SubNr)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 1, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "Normalised IOI") + scale_x_continuous(breaks=seq(1,66,1))
p_ioi_seq
```

## Exclude outlier participants
```{r outlier, echo = FALSE}
excluded_ioi <- ioi[Mean < mean(ioi$Mean)+3*mean(ioi$SD) & Mean > mean(ioi$Mean)-3*mean(ioi$SD)]
print(excluded_ioi)
```

```{r export, include = FALSE}
# export df_ioi
write.csv(df_ioi, file = "./data_ioi_onset.csv", row.names = F)
```