### 2020-04-16 material/analysis/expression/stim_d/kv.Rmd
upload analysis for kv performing condition
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis for Performing-Dynamics performances (Key Velocity)'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a rough overview of non-expressive performance from teaching-v2.0 (KV).

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
# onset
dt_kv <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv

# SubNr as a factor
dt_kv$SubNr <- as.factor(dt_kv$SubNr)

# assign RowNr
dt_kv$RowNr <- rep(1:72, nrow(dt_kv)/72)

# include only articulation
dt_kv <- dt_kv[Condition == "performing" & Skill == "dynamics"]

# remove performances which contain NA
dt_kv$Error <- 0
for (error in 1:nrow(dt_error_on)){
  dt_kv[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
}
dt_kv_ori <- dt_kv # save original data
dt_kv <- dt_kv_ori[Error == 0]
```

```{r Velocitys, include = FALSE}
# # convert bpm to ms
# dt_kv[Tempo == 120]$Tempo <- 250
# dt_kv[Tempo == 110]$Tempo <- 273
# dt_kv[Tempo == 100]$Tempo <- 300

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (duration/velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)

# assign subcomponents
dt_kv$Subcomponent <- "NA"

# for forte
for (phrase in 1:length(ls_forte2)){
  for (note in 1:length(ls_forte2[[phrase]])){
    dt_kv[RowNr == ls_forte2[[phrase]][note]]$Subcomponent <- "Forte"
  }
}

# for piano
for (phrase in 1:length(ls_piano2)){
  for (note in 1:length(ls_piano2[[phrase]])){
    dt_kv[RowNr == ls_piano2[[phrase]][note]]$Subcomponent <- "Piano"
  }
}
```

# KV
## 1. All - individual
- Average KV for each participant.

```{r kv, echo = FALSE}
# For each individual
kv <- dt_kv[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Subcomponent)]
print(kv)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kv_bar,  echo = FALSE}
p_kv <- ggplot(data = kv, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Key Velocity", subtitle = "KV: Performing-Dynamics")
p_kv
```

#### Sequence plot
- Right wrap: Subject Number

#### Forte
```{r kv_seq_for, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq_for <- dt_kv[Subcomponent == "Forte", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq_for <- ggplot(data = kv_seq_for, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
p_kv_seq_for
```

#### Piano
```{r kv_seq_pia, fig.width = 14, fig.height = 12, echo = FALSE}
# For each individual
dt_kv$TrialNr <- as.factor(dt_kv$TrialNr)
kv_seq_pia <- dt_kv[Subcomponent == "Piano", .(N = length(Velocity), Mean = mean(Velocity)), by = .(SubNr, TrialNr, RowNr)]

p_kv_seq_pia <- ggplot(data = kv_seq_pia, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Key Velocity") + scale_x_continuous(breaks=seq(1,66,1))
p_kv_seq_pia
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

```{r kv_dis_for, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis_for <- ggplot(dt_kv[Subcomponent == "Forte"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Forte)")
p_kv_dis_for
```

```{r kv_dis_pia, fig.width = 14, fig.height = 8, echo = FALSE}
p_kv_dis_pia <- ggplot(dt_kv[Subcomponent == "Piano"], aes(x = Velocity)) +
  geom_histogram(binwidth = 2, alpha = .5, position = "identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Key Velocity", y = "Count", subtitle = "KV: Performing-Dynamics(Piano)")
p_kv_dis_pia
```

## 2. All - group
- Average Velocity

```{r kv_all, echo = FALSE}
# Group mean
kv_all <- kv[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean)), by = .(Subcomponent)]
print(kv_all)
```

### Exclude outlier participants
#### Forte
```{r outlier_for, echo = FALSE}
# for forte
excluded_duration_for <- kv[Subcomponent == "Forte" & Mean < kv_all[Subcomponent == "Forte"]$Mean+3*kv_all[Subcomponent == "Forte"]$SD & Mean > kv_all[Subcomponent == "Forte"]$Mean-3*kv_all$SD]
print(excluded_duration_for)
```

#### Piano
```{r outlier_pia, echo = FALSE}
# for piano
excluded_duration_pia <- kv[Subcomponent == "Piano" & Mean < kv_all[Subcomponent == "Piano"]$Mean+3*kv_all[Subcomponent == "Piano"]$SD & Mean > kv_all[Subcomponent == "Piano"]$Mean-3*kv_all$SD]
print(excluded_duration_pia)
```

<!-- # KV Variability -->
<!-- ```{r cv, echo = FALSE} -->
<!-- # for each participant -->
<!-- cv <- dt_kv[EighthNote == "Yes", .(N = length(Velocity), CV = sd(Velocity)/mean(Velocity)) , by = .(SubNr, TrialNr)] -->
<!-- cv -->
<!-- ``` -->

<!-- ```{r cv_box, echo = FALSE} -->
<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
<!-- p_cv_box -->
<!-- ``` -->

<!-- ### Exclude outlier performances -->
<!-- ```{r cv_outlier, echo = FALSE} -->
<!-- # for all -->
<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
<!-- print(cv_all) -->

<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
<!-- print(excluded_cv) -->
<!-- ``` -->