### 2020-09-04 script/psychopy/data/summary.Rmd
summary and plot for the stimuli selection experiment
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Summary for the stimuli selection experiment"
output: html_notebook
---

Participants: Anna, Thomas and me

```{r setup, include = FALSE}
# install packages
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# setting
theme_update(legend.position = "bottom")
```

## Overall rating for each category
```{r raw, echo = FALSE}
# Create a list of data file names
lf <- list.files(".", pattern = "csv")

# Create raw_data - merge all data files into one
raw_data <- data.table()
for (i in 1:length(lf)){
  data_i <- fread(file.path(".", lf[i]), header = T, sep = ",", dec = ".")
  raw_data <- rbind(raw_data, data_i)
}
# remove unnecessary strings from midFile
raw_data$midFile <- gsub("./practice/", "", raw_data$midFile)
raw_data$midFile <- gsub("./mid/", "", raw_data$midFile)
raw_data$midFile <- gsub(".mid", "", raw_data$midFile)

# remove practice trials
data <- raw_data[expMode == "experiment"]
```

```{r summary, echo = FALSE}
data$stimCategory <- "NA"
data[startsWith(midFile, "high") == TRUE]$stimCategory <- "High"
data[startsWith(midFile, "art") == TRUE]$stimCategory <- "Art_only"
data[startsWith(midFile, "dyn") == TRUE]$stimCategory <- "Dyn_only"
data[startsWith(midFile, "low") == TRUE]$stimCategory <- "Low"

# overall rating for each category
summary_data <- data[, .(N = .N/2, meanRating = mean(rating), SD = sd(rating), SEM = sd(rating)/sqrt(.N/2)), by = .(stimCategory, ratingCategory)]
summary_data[order(ratingCategory, stimCategory)]
```

## Plot
- Error bars represent SEM.
```{r summary_plot, echo = FALSE}
plot <- ggplot(summary_data, aes(x = stimCategory, y = meanRating)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin = meanRating-SEM, ymax = meanRating + SEM), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(ratingCategory ~ .)
plot
```

## For each performance
### High performance
```{r high, echo = ALSE}
data_high <- data[stimCategory == "High", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_high$subjectNumber <- factor(data_high$subjectNumber)
data_high$Rating <- factor(data_high$rating)
data_high$midFile <- gsub("high_", "", data_high$midFile)

# plot
p_high <- ggplot(data_high, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .))
p_high
```

### Articulation-only performance
```{r art, echo = FALSE}
data_art <- data[stimCategory == "Art_only", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_art$subjectNumber <- factor(data_art$subjectNumber)
data_art$Rating <- factor(data_art$rating)
data_art$midFile <- gsub("art_", "", data_art$midFile)

# plot
p_art <- ggplot(data = data_art, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_art
```

### Dynamics-only performance
```{r dyn, echo = FALSE}
data_dyn <- data[stimCategory == "Dyn_only", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_dyn$subjectNumber <- factor(data_dyn$subjectNumber)
data_dyn$Rating <- factor(data_dyn$rating)
data_dyn$midFile <- gsub("dyn_", "", data_dyn$midFile)

# plot
p_dyn <- ggplot(data = data_dyn, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_dyn
```
### Low performance
```{r low, echo = FALSE}
data_low <- data[stimCategory == "Low", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_low$subjectNumber <- factor(data_low$subjectNumber)
data_low$Rating <- factor(data_low$rating)
data_low$midFile <- gsub("low_", "", data_low$midFile)

# plot
p_low <- ggplot(data = data_low, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_low
```

## Feedback from Anna and Thomas