### 2021-01-11 material/expstimuli/analysis/expression/stim_d/duration.Rmd
organised files and updated readme
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Duration Analysis for Performing-Dynamics performances'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a rough overview of performing performance (Duration) from teaching-v2.0.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r extract, include = FALSE}
# onset
dt_onset <- fread("../filtered/data_correct_onset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
dt_error_on <- fread("../filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".") # read an error csv
# assign RowNr
dt_onset$RowNr <- rep(1:72, nrow(dt_onset)/72)
# sort by SubNr, BlockNr, TrialNr and NoteNr
dt_onset <- dt_onset[order(SubNr, BlockNr, TrialNr, RowNr),]
# include only dynamics
dt_onset <- dt_onset[Condition == "performing" & Skill == "dynamics"]

# SubNr as a factor
dt_onset$SubNr <- as.factor(dt_onset$SubNr)
dt_onset$TrialNr <- as.factor(dt_onset$TrialNr)

# remove performances which contain NA
dt_onset$Error <- 0
for (error in 1:nrow(dt_error_on)){
  dt_onset[SubNr == dt_error_on[error]$SubNr & BlockNr == dt_error_on[error]$BlockNr & TrialNr == dt_error_on[error]$TrialNr]$Error <- 1
}
dt_onset_ori <- dt_onset # save original data
dt_onset <- dt_onset_ori[Error == 0]

# offset
dt_offset <- fread("../filtered/data_correct_offset.csv", header = T, sep = ",", dec = ".") # read a filtered csv
dt_error_off <- fread("../filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".") # read an error csv
# assign RowNr
dt_offset$RowNr <- rep(1:72, nrow(dt_offset)/72)
# sort by SubNr, BlockNr, TrialNr and NoteNr
dt_offset <- dt_offset[order(SubNr, BlockNr, TrialNr, RowNr),]
# include only dynamics
dt_offset <- dt_offset[Condition == "performing" & Skill == "dynamics"]

# SubNr as a factor
dt_offset$SubNr <- as.factor(dt_offset$SubNr)
dt_offset$TrialNr <- as.factor(dt_offset$TrialNr)

# remove performances which contain NA
dt_offset$Error <- 0
for (error in 1:nrow(dt_error_off)){
  dt_offset[SubNr == dt_error_off[error]$SubNr & BlockNr == dt_error_off[error]$BlockNr & TrialNr == dt_error_off[error]$TrialNr]$Error <- 1
}
dt_offset_ori <- dt_offset # save original data
dt_offset <- dt_offset_ori[Error == 0]
```

```{r normDu, include = FALSE}
# extract trials which have both onset and offset data
dt_onset$Eval <- "NA"
for (subnr in unique(dt_offset$SubNr)){
  for (trial in unique(dt_offset[SubNr == subnr]$Trial)){
    dt_onset[SubNr == subnr & TrialNr == trial]$Eval <- "Yes"
  }
}
dt_duration <- dt_onset[Eval == "Yes"]

# check whether each row corresponds correctly
dt_duration$Eval2 <- "NA"
for (i in 1:nrow(dt_duration)){
  if (dt_duration[i]$SubNr == dt_offset[i]$SubNr & dt_duration[i]$TrialNr == dt_offset[i]$TrialNr & dt_duration[i]$Pitch == dt_offset[i]$Pitch){
    dt_duration[i]$Eval2 <- "SAME"
  } else {
    dt_duration[i]$Eval2 <- "DIFF"
  }
}
print(unique(dt_duration$Eval2))

# TimeStamp as an integer (only offsets - I don't know why 16.04.20)
dt_offset$TimeStamp <- as.integer(dt_offset$TimeStamp)

# calculate duration
dt_duration$Duration <- dt_offset$TimeStamp - dt_duration$TimeStamp

# convert bpm to ms
dt_duration[Tempo == 120]$Tempo <- 250
dt_duration[Tempo == 110]$Tempo <- 273
dt_duration[Tempo == 100]$Tempo <- 300

# normalised Duration
dt_duration$normDu <- dt_duration$Duration/dt_duration$Tempo

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (duration/velocity)
ls_legato2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)

# assign subcomponents
dt_duration$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_forte2)){
  for (note in 1:length(ls_forte2[[phrase]])){
    dt_duration[RowNr == ls_forte2[[phrase]][note]]$Subcomponent <- "Forte"
  }
}
# Staccato
for (phrase in 1:length(ls_piano2)){
  for (note in 1:length(ls_piano2[[phrase]])){
    dt_duration[RowNr == ls_piano2[[phrase]][note]]$Subcomponent <- "Piano"
  }
}
```

# Duration
## 1. All - individual
- Average normDu for each participant

```{r duration, echo = FALSE}
# for each individual
duration <- dt_duration[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu)), by = .(SubNr, Subcomponent)]
print(duration)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r duration_bar,  echo = FALSE}
p_duration <- ggplot(data = duration, aes(x = SubNr, y = Mean)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics")
p_duration
```

#### Sequence plot
- Right wrap: Subject Number

#### All
```{r duration_seq, fig.width = 14, fig.height = 12, echo = FALSE}
# for each individual
dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
duration_seq <- dt_duration[Subcomponent != "NA", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]

p_duration_seq <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics(All)") +
  scale_x_continuous(breaks=seq(1,66,1)) + scale_y_continuous(breaks=seq(0,2,0.5), limits = c(0, 2))
p_duration_seq
```

#### Forte
```{r duration_seq_leg, fig.width = 14, fig.height = 12, echo = FALSE}
# for each individual
dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
duration_seq_leg <- dt_duration[Subcomponent == "Forte", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]

p_duration_seq_leg <- ggplot(data = duration_seq_leg, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics(Forte)") +
  scale_x_continuous(breaks=seq(1,66,1))
p_duration_seq_leg
```

#### Piano
```{r duration_seq_sta, fig.width = 14, fig.height = 12, echo = FALSE}
# for each individual
dt_duration$TrialNr <- as.factor(dt_duration$TrialNr)
duration_seq <- dt_duration[Subcomponent == "Piano", .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr, RowNr)]

p_duration_seq_sta <- ggplot(data = duration_seq, aes(x = RowNr, y = Mean, group = TrialNr, colour = TrialNr)) +
  geom_line() +
  geom_point() +
  facet_grid(SubNr ~ .) +
  labs(x = "RowNr", y = "Normalised Duration", subtitle = "Duration: Performing-Dynamics(Piano)") +
  scale_x_continuous(breaks=seq(1,66,1))
p_duration_seq_sta
```

#### Distribution plot
- Right wrap: Trial Number
- Upper wrap: Subject Number

#### Forte
```{r duration_dis_for, fig.width = 12, echo = FALSE}
p_duration_dis_for <- ggplot(dt_duration[Subcomponent == "Forte"], aes(x = normDu)) +
  geom_histogram(binwidth = .1, alpha=.5, position="identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Dynamics(Forte)")
p_duration_dis_for
```

#### Piano
```{r duration_dis_pia, fig.width = 12, echo = FALSE}
p_duration_dis_pia <- ggplot(dt_duration[Subcomponent == "Piano"], aes(x = normDu)) +
  geom_histogram(binwidth = .1, alpha = .5, position="identity") +
  facet_grid(TrialNr ~ SubNr) +
  labs(x = "Normalised Duration", y = "Count", subtitle = "Duration: Performing-Dynamics(Piano)")
p_duration_dis_pia
```

## 2. All - group
- Average normDu

```{r duration_all, echo = FALSE}
# group mean
duration_all <- duration[, .(N = length(Mean), Mean = mean(Mean), SD = sd(Mean)), by = .(Subcomponent)]
print(duration_all)
```

### Valid performances (exclude outliers)
#### Forte
```{r outlier_leg, echo = FALSE}
# for forte
excluded_duration_for <- duration[Subcomponent == "Forte" & Mean < duration_all[Subcomponent == "Forte"]$Mean+3*duration_all[Subcomponent == "Forte"]$SD & Mean > duration_all[Subcomponent == "Forte"]$Mean-3*duration_all$SD]
print(excluded_duration_for)
```
- The number of excluded performance: `r duration_all[Subcomponent == "Forte"]$N-nrow(excluded_duration_for)`

#### Piano
```{r outlier_sta, echo = FALSE}
# for piano
excluded_duration_pia <- duration[Subcomponent == "Piano" & Mean < duration_all[Subcomponent == "Piano"]$Mean+3*duration_all[Subcomponent == "Piano"]$SD & Mean > duration_all$Mean-3*duration_all[Subcomponent == "Piano"]$SD]
print(excluded_duration_pia)
```
- The number of excluded performance: `r duration_all[Subcomponent == "Piano"]$N-nrow(excluded_duration_pia)`

<!-- ```{r export, include = FALSE} -->
<!-- duration_valid <- dt_duration[, .(N = length(normDu), Mean = mean(normDu)), .(SubNr, TrialNr)] -->
<!-- fwrite(duration_valid, "./duration_valid.csv") -->
<!-- ``` -->

<!-- # Duration Variability -->
<!-- ```{r cv, echo = FALSE} -->
<!-- # for each participant -->
<!-- cv <- dt_duration[EighthNote == "Yes", .(N = length(normDu), Mean = mean(normDu), SD = sd(normDu), CV = sd(normDu)/mean(normDu)), by = .(SubNr, TrialNr)] -->
<!-- print(cv) -->
<!-- ``` -->

<!-- ```{r cv_box, echo = FALSE} -->
<!-- p_cv_box <- ggplot(cv, aes(x = SubNr, y = CV)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 0.5) -->
<!-- p_cv_box -->
<!-- ``` -->

<!-- ### Exclude outlier performacnes -->
<!-- ```{r cv_outlier, echo = FALSE} -->
<!-- # for all -->
<!-- cv_all <- cv[, .(N = length(CV), Mean = mean(CV), SD = sd(CV))] -->
<!-- print(cv_all) -->

<!-- excluded_cv <- cv[CV < cv_all$Mean+3*(cv_all$SD) & CV > cv_all$Mean-3*cv_all$SD] -->
<!-- print(excluded_cv) -->
<!-- ``` -->