### 2021-02-02 material/prerating/data/summary.Rmd
select 4 stimuli from each category
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Summary for the stimuli selection experiment"
output: html_notebook
---
- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Descriptive Stats for the generated stimuli

**Participants:** 3 including myself (author)

**Experiment brief overview:**

- Participants were asked to listen to a number of recordings and required to rate to what extent articulation/dynamics was implemented. They rated for each expression (articulation, dynamics).
- We used a 5-point likert scale for rating.
- 64 stimuli, consisting of 16 high (both articulation and dynamics implemented), 16 articulation-only (only articulation implemented), 16 dynamics-only (only dynamics implemented) and 16 low (none of them implemented).

**Technical things:**

- The stimuli were presented randomly for each participant.
- The order of rating (articulation first or dynamics first) was randomised within participants, but the half of the ratings started with articulation and the other half started with dynamics.
- There were 3 practice trials.

**Goal:**

- Pick 4 performances which represent the characteristics of each category the most.

```{r setup, include = FALSE}
# install packages
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}

# setting
theme_update(legend.position = "bottom")
```

## Overall rating for each category
- Art_only (articulation only)
- Dyn_only (dynamics only)
- High (articulation + dynamics)
- Low (none)

```{r raw, echo = FALSE}
# create raw_data - merge all data files into one
data_ls <- list.files(".", pattern = "csv")
combined <- lapply(data_ls, function(f){
  fread(paste("./", f, sep = ""), sep = ",")
})
raw_data <- do.call(rbind.data.frame, combined)

# remove unnecessary strings from midFile
raw_data$midFile <- gsub("./practice/", "", raw_data$midFile)
raw_data$midFile <- gsub("./mid/", "", raw_data$midFile)
raw_data$midFile <- gsub(".mid", "", raw_data$midFile)

# remove practice trials
data <- raw_data[expMode == "experiment"]
```

```{r summary, echo = FALSE}
data$stimCategory <- "NA"
data[startsWith(midFile, "high") == TRUE]$stimCategory <- "High"
data[startsWith(midFile, "art") == TRUE]$stimCategory <- "Art_only"
data[startsWith(midFile, "dyn") == TRUE]$stimCategory <- "Dyn_only"
data[startsWith(midFile, "low") == TRUE]$stimCategory <- "Low"

# overall rating for each category
summary_data <- data[, .(N = .N, meanRating = mean(rating), SD = sd(rating), SEM = sd(rating)/sqrt(.N)), by = .(stimCategory, ratingCategory)]
summary_data[order(ratingCategory, stimCategory)]
```

## Plot
- Error bars represent SEM.
```{r summary_plot, echo = FALSE}
plot <- ggplot(summary_data, aes(x = stimCategory, y = meanRating)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin = meanRating-SEM, ymax = meanRating + SEM), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(ratingCategory ~ .)
plot
```

## For each performance
### High performance
```{r high, echo = ALSE}
data_high <- data[stimCategory == "High", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_high$subjectNumber <- factor(data_high$subjectNumber)
data_high$Rating <- factor(data_high$rating)
data_high$midFile <- gsub("high_", "", data_high$midFile)

# plot
p_high <- ggplot(data_high, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .))
p_high
```

### Articulation-only performance
```{r art, echo = FALSE}
data_art <- data[stimCategory == "Art_only", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_art$subjectNumber <- factor(data_art$subjectNumber)
data_art$Rating <- factor(data_art$rating)
data_art$midFile <- gsub("art_", "", data_art$midFile)

# plot
p_art <- ggplot(data = data_art, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_art
```

### Dynamics-only performance
```{r dyn, echo = FALSE}
data_dyn <- data[stimCategory == "Dyn_only", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_dyn$subjectNumber <- factor(data_dyn$subjectNumber)
data_dyn$Rating <- factor(data_dyn$rating)
data_dyn$midFile <- gsub("dyn_", "", data_dyn$midFile)

# plot
p_dyn <- ggplot(data = data_dyn, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_dyn
```
### Low performance
```{r low, echo = FALSE}
data_low <- data[stimCategory == "Low", .(N = .N, rating = rating), by = .(subjectNumber, ratingCategory, midFile)]

# organise data.table
data_low$subjectNumber <- factor(data_low$subjectNumber)
data_low$Rating <- factor(data_low$rating)
data_low$midFile <- gsub("low_", "", data_low$midFile)

# plot
p_low <- ggplot(data = data_low, aes(x = midFile, y = rating, colour = subjectNumber, group = subjectNumber)) +
  geom_point(aes(shape = subjectNumber), position = position_dodge(width = .5)) +
  facet_wrap((ratingCategory ~ .)) +
  ylim(1, 5)
p_low
```

# Selecting stimuli
## High performance
- sum the ratings for both articulation and dynamics and rank the performances by Sum

```{r ranking-high, echo = FALSE}
data_high <- data_high[order(midFile, ratingCategory)]
data_high_sep <- data_high[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
data_high_both <- data_high[, .(N = .N, Sum = sum(rating)), by = .(midFile)]

data_high_both[order(-Sum)]
```

### Selected 4 performances
```{r, echo = FALSE}
select_high <- data_high_both[order(-Sum)]
select_high$Rank <- 1:nrow(select_high)
select_high$Category <- "High"

# dummy
select_high$Sample <- 0
select_high$Diff <- 0
select_high$Total <- 0

select_high[Rank < 5]
```

## Articulation performance
- sum the ratings for articulation (Sum) and calculate the difference between articulation and dynamics ratings (Diff)
- find bigger articulation ratings and bigger differences between articulation and dynamics ratings
- rank the performances by Total

```{r ranking-art, echo = FALSE}
data_art <- data_art[order(midFile, ratingCategory)]
data_art_sep <- data_art[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
data_art_art <- data_art[ratingCategory == "articulation", .(N = .N, Sum = sum(rating)), by = .(midFile)]
data_art_art$Diff <- data_art_sep[ratingCategory == "articulation"]$Sum - data_art_sep[ratingCategory == "dynamics"]$Sum

# articulation ratings + difference between articulation and dynamics ratings (bigger values are good stimuli)
data_art_art$Total <- data_art_art$Sum + data_art_art$Diff
data_art_art[order(-Total)]
```

### Selected 4 performances
```{r, echo = FALSE}
select_art <- data_art_art[order(-Total)]
select_art$Rank <- 1:nrow(select_art)
select_art$Category <- "Art_only"

# randomly select 3 from Total == 22
select_art$Sample <- 0
select_art[Rank == 1]$Sample <- 1
select_art[Total == 22]$Sample <- sample(c(1:nrow(select_art[Total == 22])), replace = FALSE)
select_art[Sample != 0 & Sample < 4]
```

## Dynamics performance
- sum the ratings for dynamics (Sum) and calculate the difference between dynamics and articulation ratings (Diff)
- find bigger dynamics ratings and bigger differences between dynamics and articulation ratings
- rank the performances by Total

```{r ranking-dyn, echo = FALSE}
data_dyn <- data_dyn[order(midFile, ratingCategory)]
data_dyn_sep <- data_dyn[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
data_dyn_dyn <- data_dyn[ratingCategory == "dynamics", .(N = .N, Sum = sum(rating)), by = .(midFile)]
data_dyn_dyn$Diff <- data_dyn_sep[ratingCategory == "dynamics"]$Sum - data_dyn_sep[ratingCategory == "articulation"]$Sum

# dynamics ratings + difference between dynamics and articulation ratings (bigger values are good stimuli)
data_dyn_dyn$Total <- data_dyn_dyn$Sum + data_dyn_dyn$Diff
data_dyn_dyn[order(-Total)]
```

### Selected 4 performances
```{r, echo = FALSE}
select_dyn <- data_dyn_dyn[order(-Total)]
select_dyn$Rank <- 1:nrow(select_dyn)
select_dyn$Category <- "Dyn_only"

# dummy
select_dyn$Sample <- 0

select_dyn[Total >= 11]
```

## Low performance
- sum the ratings for both articulation and dynamics and rank the performances by Sum

```{r ranking-low, echo = FALSE}
data_low <- data_low[order(midFile, ratingCategory)]
data_low_sep <- data_low[, .(N = .N, Sum = sum(rating)), by = .(midFile, ratingCategory)]
data_low_both <- data_low[, .(N = .N, Sum = sum(rating)), by = .(midFile)]

data_low_both[order(Sum)]
```

### Selected 4 performances
```{r, echo = FALSE}
select_low <- data_low_both[order(Sum)]
select_low$Rank <- 1:nrow(select_low)
select_low$Category <- "Low"
select_low <- select_low[Sum == min(select_low$Sum)]

# randomly select 4
select_low$Sample <- sample(c(1:nrow(select_low)), replace = FALSE)

# dummy
select_low$Diff <- 0
select_low$Total <- 0

select_low[Sample < 5]
```
# Export selected stimuli
```{r, echo = FALSE}
selected <- rbind(select_high[Rank < 5], select_art[Sample != 0 & Sample < 4], select_dyn[Total >= 11], select_low[Sample < 5])
selected

# export txt file
filename = paste("./selected_", format(Sys.time(), "%s-%d%m%y"), ".txt", sep = "") # current time
fwrite(selected, filename)
```