### 2021-09-16 material/expstimuli/audio/tweaked/analysis.Rmd

---
title: "Music Adaptation: Analysis for selected stimuli"
output: html_notebook
---

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r midi, include = FALSE}
# create a list of data file names
lf <- list.files("../selected/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../selected/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:144, nrow(data)/144)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$SubArt <- "NA"
data$SubDyn <- "NA"
data$MidFileShort <- gsub("art_only", "a", data$MidFile)
data$MidFileShort <- gsub("dyn_only", "d", data$MidFileShort)
data$MidFileShort <- gsub("high", "h", data$MidFileShort)
data$MidFileShort <- gsub("low", "l", data$MidFileShort)

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (velocity)
ls_legato_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

# IOI
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Interval == number]$SubArt <- "LtoS"
 dt_ioi[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_ioi[Interval == number]$SubArt <- "StoL"
 dt_ioi[Interval == number]$SubDyn <- "PtoF"
}
```

- There are 4 base tempi.

## Histogram
```{r ioi-distribution, echo = FALSE}
gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean
- The red line indicates the ideal tempo (100bpm - IOI: 300 ms)

```{r ioi-mean, echo = FALSE}
# mean
dt_ioi[SubArt != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), "MidFile"]

ggboxplot(dt_ioi[SubArt != "NA"], x = "MidFileShort", y = "IOI", add = "jitter", width = 0.8) + geom_hline(yintercept = 300, linetype = "dashed", color = "red")
```

# KOT
- For art_only and high stimuli, there should be a bimodal distribution to represent legato and staccato.

- For dyn_only and low stimuli, there should be a normal distribution.

```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time[row] - data_onset$time[row+1] # offset(n) - onset(n+1)
   }
}

dt_kot <- data_onset[NoteNr != 1]
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_kot[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_kot[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_kot[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_kot[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_kot[Interval == number]$SubArt <- "LtoS"
 dt_kot[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_kot[Interval == number]$SubArt <- "StoL"
 dt_kot[Interval == number]$SubDyn <- "PtoF"
}
```

## Histogram
```{r kot-distribution, echo = FALSE}
gghistogram(dt_kot[SubArt != "NA"], x = "KOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean
```{r kot-mean, echo = FALSE}
# mean
dt_kot[SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, SubArt)]

ggboxplot(dt_kot[SubArt != "NA"], x = "MidFileShort", y = "KOT", color = "SubArt", add = "jitter", width = 0.8)
```

## Tweak
- For dyn_only and low stimuli, it seems the KOT values are too close to the staccato range. We calculated the difference between the mean of each stimulus and the grand mean of art_only and high stimuli and add that difference to all KOT values for dyn_only and low, so that the KOT values should shift closer to the grand mean of art_only and high stimuli.

```{r kot-tweak, echo = FALSE}
# grandMean of art_only & high stimuli
kot_grandMean <- mean(dt_kot[grepl("art_only", MidFile) | grepl("high", MidFile) & SubArt != "NA"]$KOT)

# calculate the diff between each KOT mean and grandMean
tweak_data <- dt_kot[grepl("dyn_only", MidFile) | grepl("low", MidFile) & SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]
tweak_data$Diff <- round(abs(tweak_data$Mean-kot_grandMean), 0)
tweak_data

# add each fixed Diff value to TimeStamp
dt_kot$FixedKOT <- dt_kot$KOT
for (midfile in unique(tweak_data$MidFile)){
   dt_kot[MidFile == midfile]$FixedKOT <- dt_kot[MidFile == midfile]$KOT+tweak_data[MidFile == midfile]$Diff
}
```

d KOT)
```{r kot-distribution-fixed, echo = FALSE}
gghistogram(dt_kot[SubArt != "NA"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean (Fixed KOT)
```{r kot-mean-fixed, echo = FALSE}
# mean
dt_kot[SubArt != "NA", .(N = .N, Mean = mean(FixedKOT), SD = sd(FixedKOT)), by = .(MidFile, SubArt)]

ggboxplot(dt_kot[SubArt != "NA"], x = "MidFileShort", y = "FixedKOT", color = "SubArt", add = "jitter", width = 0.8)
```

# KV

```{r vel, include = FALSE}
dt_vel <- data_onset
dt_vel$NoteOnsetNr <- rep(1:72, nrow(dt_vel)/72)
# assign Subcomponents
# for each note
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_legato_2[[phrase]][note]]$SubArt <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_staccato_2[[phrase]][note]]$SubArt <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_forte_2[[phrase]][note]]$SubDyn <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_piano_2[[phrase]][note]]$SubDyn <- "Piano"
   }
}
```

## Histogram
```{r vel-distribution, echo = FALSE}
gghistogram(dt_vel[SubArt != "NA"], x = "Velocity", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
```

## Mean
```{r vel-mean, echo = FALSE}
# mean
dt_vel[SubDyn != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(MidFile, SubDyn)]

ggboxplot(dt_vel[SubArt != "NA"], x = "MidFileShort", y = "Velocity", color = "SubDyn", add = "jitter", width = 0.8)
```

# KV Diff
```{r vel-diff, include = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_vel_diff[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_vel_diff[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_vel_diff[Interval == number]$SubArt <- "LtoS"
 dt_vel_diff[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_vel_diff[Interval == number]$SubArt <- "StoL"
 dt_vel_diff[Interval == number]$SubDyn <- "PtoF"
}
```

## Histogram
```{r vel-diff-distribution, echo = FALSE}
gghistogram(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "Diff", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
```

## Mean
```{r vel-diff-mean, echo = FALSE}
# mean
dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(MidFile, SubDyn)]

ggboxplot(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "MidFileShort", y = "Diff", color = "SubDyn", add = "jitter", width = 0.8)
```

# Export stimuli with fixed KOT as mid files
```{r export, include = FALSE}
# remove unnecessary columns
data_onset$IOI <- NULL
data_onset$KOT <- NULL
data_onset$Diff <- NULL
# fix time for onset and offset data
data_onset$timeFixed <- data_onset$time
data_offset$timeFixed <- data_offset$time
for (midfile in unique(tweak_data$MidFile)){
   # offset[i] = onset[i+1]+KOT[i]
   for (i in 1:nrow(data_onset[MidFile == midfile])-1){
      data_offset[MidFile == midfile][i]$timeFixed <- data_onset[MidFile == midfile][i+1]$time+dt_kot[MidFile == midfile][i]$FixedKOT
   }
}

# export each instance as txt file
for (stim in unique(data_onset$MidFile)){
   onset <- data_onset[MidFile == stim]
   offset <- data_offset[MidFile == stim]
   instance <- rbind(onset, offset)
   # Note_On/Off >> 1 or 0
   instance$KeyOnOff <- 0
   instance[event == "Note On"]$KeyOnOff <- 1
   # sort by timeFixed
   instance <- instance[order(timeFixed)]

   filename = paste(stim, ".txt", sep = "")
   fwrite(instance, filename)
}
```

```{bash, echo = TRUE}
# python
source /Users/atsukotominaga/Pyenv/bin/activate
python mid_export.py
```