### 2021-10-10 analysis/filtering.Rmd
add basic analysis scripts
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Preprocessing: filtering (experiment)"
output: html_notebook
---

```{r setup, include = FALSE}
# set working directory
if (!require("here")) {install.packages("here"); require("here")}
here::i_am("filtering.Rmd")

# create a folder if not exists
if (!file.exists(here("filtered"))){
  dir.create(here("filtered"))
}

# read functions
source(here("function.R"))

# read a text file for ideal performance
dt_ideal <- read.table(here("ideal.txt"))
colnames(dt_ideal) <- "Pitch"
dt_ideal$RowNr <- 1:nrow(dt_ideal)
setcolorder(dt_ideal, c(2, 1))

# create a list of data file names
lf <- list.files(here("raw_data"), pattern = "txt")

# create raw_data - merge all data files into one
raw_data <- data.table()
for (i in 1:length(lf)){
  data_i <- fread(file.path(here("raw_data"), lf[i]), header = F, sep = " ", dec = ".")
  raw_data <- rbind(raw_data, data_i)
}

# add column namesls
colnames(raw_data) <- c("NoteNr", "TimeStamp", "Pitch", "Velocity", "Key_OnOff", "Device", "SubNr", "TrialNr", "Stimuli", "Session")

# clean raw_data
raw_data$NoteNr <- as.numeric(gsub(",", "", raw_data$NoteNr))
raw_data$Session <- gsub(";", "", raw_data$Session)

# sort by SubNr, TrialNr
raw_data <- raw_data[order(raw_data$SubNr, raw_data$TrialNr),]

# raw_data without metronome
dt_all <- raw_data[Pitch != 31 & Pitch != 34]

# onset and offset for experimental session
dt_onset <- dt_all[Key_OnOff == 1 & Session == "experiment"]
dt_offset <- dt_all[Key_OnOff == 0 & Session == "experiment"]
```

# Onset
## Detect erroneous trials
```{r, echo = FALSE}
dt_error_onset <- checker(dt_onset, dt_ideal)

# mark imperfect performances
dt_onset$Error <- 0
for (error in 1:nrow(dt_error_onset)){
  dt_onset[SubNr == dt_error_onset$SubNr[error] & TrialNr == dt_error_onset$TrialNr[error]]$Error <- 1
}

dt_correct_onset_1 <- dt_onset[Error == 0]
dt_correct_onset_1$RowNr <- rep(1:72, nrow(dt_correct_onset_1)/72)
setcolorder(dt_correct_onset_1, c(12, 1:11))

# export csv
fwrite(dt_correct_onset_1, file = here("filtered", "dt_correct_onset_1.txt"))

# print the result
dt_error_onset
```

# Offset

## Detect erroneous trials
```{r, echo = FALSE}
dt_error_offset <- checker(dt_offset, dt_ideal)

# mark imperfect performances
dt_offset$Error <- 0
for (error in 1:nrow(dt_error_offset)){
  dt_offset[SubNr == dt_error_offset$SubNr[error] & TrialNr == dt_error_offset$TrialNr[error]]$Error <- 1
}

dt_correct_offset_1 <- dt_offset[Error == 0]
dt_correct_offset_1$RowNr <- rep(1:72, nrow(dt_correct_offset_1)/72)
setcolorder(dt_correct_offset_1, c(12, 1:11))

# export csv
fwrite(dt_correct_offset_1, file = here("filtered", "dt_correct_offset_1.txt"))

# print the result
dt_error_offset
```

# Checking