### 2021-12-02 analysis/primary.Rmd
add kot analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Primary analysis'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of primary analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline
- IOIs
- KOT
- KV
- KV-Diff

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r file, include = FALSE}
onset = "filtered/dt_correct_onset_1.txt"
offset = "filtered/dt_correct_offset_1.txt"
```

```{r prep, include = FALSE}
data_on <- fread(onset, header = T, sep = ",", dec = ".") # read a trimmed csv
data_off <- fread(offset, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
data_on$SubNr <- as.factor(data_on$SubNr)

# assign stimuli categories
data_on$Stimuli <- as.numeric(gsub(".mid", "", data_on$Stimuli))
data_on$Category <- "NA"
data_on[Stimuli < 5]$Category <- "art_only"
data_on[Stimuli > 4 & Stimuli < 9]$Category <- "dyn_only"
data_on[Stimuli > 8 & Stimuli < 13]$Category <- "high"
data_on[Stimuli > 12]$Category <- "low"

# assign articulation (yes/no) & dynamics (yes/no)
data_on$Articulation <- "NA"
data_on$Dynamics <- "NA"
data_on[Category == "art_only"]$Articulation <- "Present"
data_on[Category == "art_only"]$Dynamics <- "Absent"
data_on[Category == "dyn_only"]$Articulation <- "Absent"
data_on[Category == "dyn_only"]$Dynamics <- "Present"
data_on[Category == "high"]$Articulation <- "Present"
data_on[Category == "high"]$Dynamics <- "Present"
data_on[Category == "low"]$Articulation <- "Absent"
data_on[Category == "low"]$Dynamics <- "Absent"

# IOIs
dt_ioi <- data_on
dt_ioi$IOI <- diff(c(0, dt_ioi$TimeStamp))

# remove the first note
dt_ioi <- dt_ioi[RowNr != 1]

# assign Interval
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# define Subcomponents
# for each interval
ls_l_f <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_s_p <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
change_1 <- c(5, 17, 47) # legato to staccato / forte to piano
change_2 <- c(8, 20, 39) # staccato to legato / piano to forte

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_ioi[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_ioi[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_ioi[Interval == number]$Subcomponent <- "LFtoPS"
}
for (number in change_2){
  dt_ioi[Interval == number]$Subcomponent <- "SPtoLF"
}

# KOT
onset_trials <- data_on[, .(N = .N), by = .(SubNr, TrialNr)]
offset_trials <- data_off[, .(N = .N), by = .(SubNr, TrialNr)]
valid_trials <- rbind(onset_trials, offset_trials)
valid_trials$Duplicate <- duplicated(valid_trials)
valid_trials_included <- valid_trials[Duplicate == TRUE]

dt_kot_on <- data.table()
for (row in 1:nrow(valid_trials_included)){
  current <- data_on[SubNr == valid_trials_included$SubNr[row] & TrialNr == valid_trials_included$TrialNr[row]]
  dt_kot_on <- rbind(dt_kot_on, current)
}

dt_kot_off <- data.table()
for (row in 1:nrow(valid_trials_included)){
  current <- data_off[SubNr == valid_trials_included$SubNr[row] & TrialNr == valid_trials_included$TrialNr[row]]
  dt_kot_off <- rbind(dt_kot_off, current)
}

# sort by SubNr, TrialNr and RowNr
dt_kot_on <- dt_kot_on[order(SubNr, TrialNr, RowNr)]
dt_kot_off <- dt_kot_off[order(SubNr, TrialNr, RowNr)]

# Offset 1 - Onset 2
dt_kot_on$KOT <- NA
for (row in 1:nrow(dt_kot_on)){
  if (row < nrow(dt_kot_on)){
    dt_kot_on$KOT[row+1] <- dt_kot_off$TimeStamp[row] - dt_kot_on$TimeStamp[row+1] # offset(n) - onset(n+1)
  }
}

# remove the first note
dt_kot_on <- dt_kot_on[RowNr != 1]

# assign Interval
dt_kot_on$Interval <- rep(1:71, nrow(dt_kot_on)/71)

# assign Subcomponents
dt_kot_on$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_kot_on[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_kot_on[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_kot_on[Interval == number]$Subcomponent <- "LFtoPS"
}
for (number in change_2){
  dt_kot_on[Interval == number]$Subcomponent <- "SPtoLF"
}
# KV
dt_vel <- data_on
dt_vel$Diff <- diff(c(0, dt_vel$Velocity))

# remove the first note
dt_vel_diff <- dt_vel[RowNr != 1]
dt_vel$Diff <- NULL # remove Diff from dt_vel

# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# for each note (velocity)
ls_l_f_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_s_p_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# assign Subcomponents
# for each note
dt_vel$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f_2)){
  for (note in 1:length(ls_l_f_2[[phrase]])){
    dt_vel[RowNr == ls_l_f_2[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p_2)){
  for (note in 1:length(ls_s_p_2[[phrase]])){
    dt_vel[RowNr == ls_s_p_2[[phrase]][note]]$Subcomponent <- "S-P"
  }
}

dt_vel_diff$Subcomponent <- "NA"
# legato & forte
for (phrase in 1:length(ls_l_f)){
  for (note in 1:length(ls_l_f[[phrase]])){
    dt_vel_diff[Interval == ls_l_f[[phrase]][note]]$Subcomponent <- "L-F"
  }
}
# staccato & piano
for (phrase in 1:length(ls_s_p)){
  for (note in 1:length(ls_s_p[[phrase]])){
    dt_vel_diff[Interval == ls_s_p[[phrase]][note]]$Subcomponent <- "S-P"
  }
}
for (number in change_1){
  dt_vel_diff[Interval == number]$Subcomponent <- "LFtoSP"
}
for (number in change_2){
  dt_vel_diff[Interval == number]$Subcomponent <- "SPtoLF"
}
```

# IOI 
## 1. Individual

```{r ioi, echo = FALSE}
# for each individual
ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Stimuli, Articulation, Dynamics)]
ioi_ind <- ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)]
ioi_ind
```

### Plot
#### Boxplot

```{r ioi-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(ioi[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

## 2. All - group

```{r ioi-all, echo = FALSE}
# group mean
ioi_all <- ioi_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics)]
ioi_all
```

### Plot
#### Box plot

```{r ioi-all-plot, echo = FALSE}
ggboxplot(ioi_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "IOIs (ms)", title = "IOI")
```

# Articulation
## 1. Interval
```{r kot, echo = FALSE}
# for each interval
kot <- dt_kot_on[Subcomponent == "L-F" | Subcomponent == "S-P", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
kot_ind <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Subcomponent)]
kot_ind
```

### Plot
#### Boxplot

```{r kot-plot,  echo = FALSE, fig.width = 4.5}
ggboxplot(kot[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
ggboxplot(kot[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", facet.by = "SubNr", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

## 2. All - group
```{r kot-all, echo = FALSE}
# group mean
kot_all <- kot_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
kot_all
```

### Plot
#### Boxplot

```{r kot-all-plot, echo = FALSE}
ggboxplot(kot_ind[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Legato)")
ggboxplot(kot_ind[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Articulation", "Mean", color = "Dynamics", xlab = "Articulation", ylab = "KOT (ms)", title = "KOT (Staccato)")
```

# Dynamics
## 1. Individual
```{r vel, echo = FALSE}
# for each note
vel <- dt_vel[Subcomponent != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel_ind <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Subcomponent)]
vel_ind
```

### Plot
#### Boxplot

```{r vel-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(vel[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Forte)")
ggboxplot(vel[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "SubNr", ylab = "Velocity", title = "KV (Piano)")
```

## 2. All - group
```{r vel-all, echo = FALSE}
# group mean
vel_all <- vel_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
vel_all
```

### Plot
#### Boxplot

```{r vel-all-plot,  echo = FALSE}
ggboxplot(vel_ind[Subcomponent == "L-F", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Forte)")
ggboxplot(vel_ind[Subcomponent == "S-P", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity", title = "KV (Piano)")
```

# KV-Diff
## 1. Individual

```{r vel-diff, echo = FALSE}
# for each individual
vel_diff <- dt_vel_diff[Subcomponent == "LFtoSP" | Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Stimuli, Articulation, Dynamics, Subcomponent)]
vel_diff_ind <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics, Subcomponent)]
vel_diff_ind
```

### Plot
#### Boxplot

```{r vel-diff-plot, echo = FALSE, fig.width = 4.5}
ggboxplot(vel_diff[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (LFtoSP)")
ggboxplot(vel_diff[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Stimuli, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", facet.by = "SubNr", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV-Diff (SPtoLF)")
```

## 2. All - group
```{r vel-diff-all, echo = FALSE}
# group mean
vel_diff_all <- vel_diff_ind[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Articulation, Dynamics, Subcomponent)]
vel_diff_all
```

### Plot
#### Boxplot

```{r vel-diff-all-plot,  echo = FALSE}
ggboxplot(vel_diff_ind[Subcomponent == "LFtoSP", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (LF to SP)")
ggboxplot(vel_diff_ind[Subcomponent == "SPtoLF", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Articulation, Dynamics)], "Dynamics", "Mean", color = "Articulation", xlab = "Dynamics", ylab = "Velocity Difference", title = "KV (SP to LF)")
```