### 2023-01-17 material/expstimuli/audio/tweaked/analysis.Rmd
update figures
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Adaptation: Analysis for selected stimuli"
output: html_notebook
---

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

```{r midi, include = FALSE}
# create a list of data file names
lf <- list.files("../selected/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../selected/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:144, nrow(data)/144)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$SubArt <- "NA"
data$SubDyn <- "NA"
data$MidFileShort <- gsub("art_only", "a", data$MidFile)
data$MidFileShort <- gsub("dyn_only", "d", data$MidFileShort)
data$MidFileShort <- gsub("high", "h", data$MidFileShort)
data$MidFileShort <- gsub("low", "l", data$MidFileShort)

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (velocity)
ls_legato_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

# IOI
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Interval == number]$SubArt <- "LtoS"
 dt_ioi[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_ioi[Interval == number]$SubArt <- "StoL"
 dt_ioi[Interval == number]$SubDyn <- "PtoF"
}
```

- There are 4 base tempi.

## Histogram
```{r ioi-distribution, echo = FALSE}
# change labelling
dt_ioi[MidFile == "high_1"]$MidFile <- "both_1"
dt_ioi[MidFile == "high_2"]$MidFile <- "both_2"
dt_ioi[MidFile == "high_3"]$MidFile <- "both_3"
dt_ioi[MidFile == "high_4"]$MidFile <- "both_4"
dt_ioi[MidFile == "low_1"]$MidFile <- "none_1"
dt_ioi[MidFile == "low_2"]$MidFile <- "none_2"
dt_ioi[MidFile == "low_3"]$MidFile <- "none_3"
dt_ioi[MidFile == "low_4"]$MidFile <- "none_4"

gghistogram(dt_ioi[SubArt != "NA"], x = "IOI", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean
- The red line indicates the ideal tempo (100bpm - IOI: 300 ms)

```{r ioi-mean, echo = FALSE}
# mean
dt_ioi[SubArt != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), "MidFile"]

ggboxplot(dt_ioi[SubArt != "NA"], x = "MidFileShort", y = "IOI", add = "jitter", width = 0.8) + geom_hline(yintercept = 300, linetype = "dashed", color = "red")
```

# KOT
```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time[row] - data_onset$time[row+1] # offset(n) - onset(n+1)
   }
}

dt_kot <- data_onset[NoteNr != 1]
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_kot[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_kot[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_kot[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_kot[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_kot[Interval == number]$SubArt <- "LtoS"
 dt_kot[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_kot[Interval == number]$SubArt <- "StoL"
 dt_kot[Interval == number]$SubDyn <- "PtoF"
}
```

## Histogram

- For art_only and both stimuli, there should be a bimodal distribution to represent the implementation of legato and staccato.

- For dyn_only and none stimuli, there should be a normal distribution.

```{r kot-distribution, echo = FALSE}
# change labelling
dt_kot[MidFile == "high_1"]$MidFile <- "both_1"
dt_kot[MidFile == "high_2"]$MidFile <- "both_2"
dt_kot[MidFile == "high_3"]$MidFile <- "both_3"
dt_kot[MidFile == "high_4"]$MidFile <- "both_4"
dt_kot[MidFile == "low_1"]$MidFile <- "none_1"
dt_kot[MidFile == "low_2"]$MidFile <- "none_2"
dt_kot[MidFile == "low_3"]$MidFile <- "none_3"
dt_kot[MidFile == "low_4"]$MidFile <- "none_4"

gghistogram(dt_kot[SubArt != "NA" & SubArt != "LtoS" & SubArt != "StoL"], x = "KOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean
```{r kot-mean, echo = FALSE}
# mean
dt_kot[SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, SubArt)]

ggboxplot(dt_kot[SubArt != "NA"], x = "MidFileShort", y = "KOT", color = "SubArt", add = "jitter", width = 0.8)
```

## Tweak
- For dyn_only and none stimuli, it seems the KOT values are too close to the staccato range. We calculated the difference between the mean of each stimulus and the grand mean of art_only and both stimuli and add that difference to all KOT values for dyn_only and none stimuli, so that the KOT values should shift closer to the grand mean of art_only and both stimuli.

## Histogram (Fixed KOT)
```{r kot-tweak, echo = FALSE}
# grandMean of art_only & both stimuli
kot_grandMean <- mean(dt_kot[grepl("art_only", MidFile) | grepl("both", MidFile) & SubArt != "NA"]$KOT)

# calculate the diff between each KOT mean and grandMean
tweak_data <- dt_kot[grepl("dyn_only", MidFile) | grepl("none", MidFile) & SubArt != "NA", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile)]
tweak_data$Diff <- round(abs(tweak_data$Mean-kot_grandMean), 0)
tweak_data

# add each fixed Diff value to TimeStamp
dt_kot$FixedKOT <- dt_kot$KOT
for (midfile in unique(tweak_data$MidFile)){
   dt_kot[MidFile == midfile]$FixedKOT <- dt_kot[MidFile == midfile]$KOT+tweak_data[MidFile == midfile]$Diff
}
```

```{r kot-distribution-fixed, echo = FALSE}
gghistogram(dt_kot[SubArt != "NA" & SubArt != "LtoS" & SubArt != "StoL"], x = "FixedKOT", color = "SubArt", facet.by = "MidFile", bins = 50, rug = TRUE)
```

## Mean (Fixed KOT)
```{r kot-mean-fixed, echo = FALSE}
# mean
dt_kot[SubArt != "NA", .(N = .N, Mean = mean(FixedKOT), SD = sd(FixedKOT)), by = .(MidFile, SubArt)]

ggboxplot(dt_kot[SubArt != "NA"], x = "MidFileShort", y = "FixedKOT", color = "SubArt", add = "jitter", width = 0.8)
```

# KV
```{r vel, include = FALSE}
dt_vel <- data_onset
dt_vel$NoteOnsetNr <- rep(1:72, nrow(dt_vel)/72)
# assign Subcomponents
# for each note
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_legato_2[[phrase]][note]]$SubArt <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_staccato_2[[phrase]][note]]$SubArt <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_forte_2[[phrase]][note]]$SubDyn <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel[NoteOnsetNr == ls_piano_2[[phrase]][note]]$SubDyn <- "Piano"
   }
}
```

## Histogram

- For dyn_only and both stimuli, there should be a bimodal distribution to represent the implementation of forte and piano.

- For art_only and none stimuli, there should be a normal distribution.

```{r vel-distribution, echo = FALSE}
# change labelling
dt_vel[MidFile == "high_1"]$MidFile <- "both_1"
dt_vel[MidFile == "high_2"]$MidFile <- "both_2"
dt_vel[MidFile == "high_3"]$MidFile <- "both_3"
dt_vel[MidFile == "high_4"]$MidFile <- "both_4"
dt_vel[MidFile == "low_1"]$MidFile <- "none_1"
dt_vel[MidFile == "low_2"]$MidFile <- "none_2"
dt_vel[MidFile == "low_3"]$MidFile <- "none_3"
dt_vel[MidFile == "low_4"]$MidFile <- "none_4"

gghistogram(dt_vel[SubArt != "NA"], x = "Velocity", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
```

## Mean
```{r vel-mean, echo = FALSE}
# mean
dt_vel[SubDyn != "NA", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(MidFile, SubDyn)]

ggboxplot(dt_vel[SubArt != "NA"], x = "MidFileShort", y = "Velocity", color = "SubDyn", add = "jitter", width = 0.8)
```

# KV Diff
```{r vel-diff, include = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# assign Subcomponents
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_vel_diff[Interval == ls_legato[[phrase]][note]]$SubArt <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$SubArt <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_vel_diff[Interval == ls_legato[[phrase]][note]]$SubDyn <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_vel_diff[Interval == ls_staccato[[phrase]][note]]$SubDyn <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_vel_diff[Interval == number]$SubArt <- "LtoS"
 dt_vel_diff[Interval == number]$SubDyn <- "FtoP"
}
for (number in change_2){
 dt_vel_diff[Interval == number]$SubArt <- "StoL"
 dt_vel_diff[Interval == number]$SubDyn <- "PtoF"
}
```

## Histogram

- For dyn_only and both stimuli, there should be a bimodal distribution to represent the transition from forte to piano or from piano to forte.

- For art_only and none stimuli, there should be a normal distribution.

```{r vel-diff-distribution, echo = FALSE}
# change labelling
dt_vel_diff[MidFile == "high_1"]$MidFile <- "both_1"
dt_vel_diff[MidFile == "high_2"]$MidFile <- "both_2"
dt_vel_diff[MidFile == "high_3"]$MidFile <- "both_3"
dt_vel_diff[MidFile == "high_4"]$MidFile <- "both_4"
dt_vel_diff[MidFile == "low_1"]$MidFile <- "none_1"
dt_vel_diff[MidFile == "low_2"]$MidFile <- "none_2"
dt_vel_diff[MidFile == "low_3"]$MidFile <- "none_3"
dt_vel_diff[MidFile == "low_4"]$MidFile <- "none_4"

gghistogram(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "Diff", facet.by = "MidFile", color = "SubDyn", bins = 50, rug = TRUE)
```

## Mean
```{r vel-diff-mean, echo = FALSE}
# mean
dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(MidFile, SubDyn)]

ggboxplot(dt_vel_diff[SubDyn == "FtoP" | SubDyn == "PtoF"], x = "MidFileShort", y = "Diff", color = "SubDyn", add = "jitter", width = 0.8)
```

```{r export, include = FALSE}
knitr::purl("analysis.Rmd")
```

<!-- # Export stimuli with fixed KOT as mid files -->
<!-- ```{r export, include = FALSE} -->
<!-- # remove unnecessary columns -->
<!-- data_onset$IOI <- NULL -->
<!-- data_onset$KOT <- NULL -->
<!-- data_onset$Diff <- NULL -->
<!-- # fix time for onset and offset data -->
<!-- data_onset$timeFixed <- data_onset$time -->
<!-- data_offset$timeFixed <- data_offset$time -->
<!-- for (midfile in unique(tweak_data$MidFile)){ -->
<!--    # offset[i] = onset[i+1]+KOT[i] -->
<!--    for (i in 1:nrow(data_onset[MidFile == midfile])-1){ -->
<!--       data_offset[MidFile == midfile][i]$timeFixed <- data_onset[MidFile == midfile][i+1]$time+dt_kot[MidFile == midfile][i]$FixedKOT -->
<!--    } -->
<!-- } -->

<!-- # export each instance as txt file -->
<!-- for (stim in unique(data_onset$MidFile)){ -->
<!--    onset <- data_onset[MidFile == stim] -->
<!--    offset <- data_offset[MidFile == stim] -->
<!--    instance <- rbind(onset, offset) -->
<!--    # Note_On/Off >> 1 or 0 -->
<!--    instance$KeyOnOff <- 0 -->
<!--    instance[event == "Note On"]$KeyOnOff <- 1 -->
<!--    # sort by timeFixed -->
<!--    instance <- instance[order(timeFixed)] -->

<!--    filename = paste(stim, ".txt", sep = "") -->
<!--    fwrite(instance, filename) -->
<!-- } -->
<!-- ``` -->

<!-- ```{bash, echo = TRUE} -->
<!-- # python -->
<!-- source /Users/atsukotominaga/Pyenv/bin/activate -->
<!-- python mid_export.py -->
<!-- ``` -->